---
title: "Summary_Statistics"
author: "Veronica Leary" "Brandon Yee"
date: "`r Sys.Date()`"
output: pdf_document
---
---

# Introduction

This report generates and displays summary statistics of text message counts by **Group** and **Time Point**.  
We calculate measures of central tendency and variability to understand texting patterns across two groups and over time.

---

# Load Required Libraries

```{r setup, message=FALSE, warning=FALSE}
library(dplyr)      # Data manipulation
library(tidyr)      # Reshaping data
library(readr)      # Reading CSV files
library(gt)         # Generating clean, styled summary tables
```

---

# Data Loading and Cleaning

```{r load-data}
# Read in the dataset and clean header rows
data <- read.csv("TextMessages.csv")  # Reads the dataset into an R dataframe

# Reshape from wide to long format
data_long <- data %>%
  mutate(across(c(Baseline, Six_months), as.numeric),
         Group = as.factor(Group)) %>%
  pivot_longer(cols = c(Baseline, Six_months),
               names_to = "Time",
               values_to = "TextMessages")
```

---

# Summary Statistics Calculation

We calculate:

- **Count** of observations per group/time
- **Mean**, **Median**, and **Standard Deviation** of text message counts

```{r summary-table}
summary_table <- data_long %>%
  group_by(Group, Time) %>%
  summarise(
    Count = n(),
    Mean = round(mean(TextMessages, na.rm = TRUE), 2),
    Median = round(median(TextMessages, na.rm = TRUE), 2),
    SD = round(sd(TextMessages, na.rm = TRUE), 2),
    .groups = "drop"
  )
```

---

# Summary Table Output

```{r summary-display}
summary_table %>%
  gt() %>%
  tab_header(
    title = "Summary Statistics of Text Messages",
    subtitle = "Grouped by Treatment Group and Time Point"
  ) %>%
  cols_label(
    Group = "Group",
    Time = "Time Point",
    Count = "N",
    Mean = "Mean",
    Median = "Median",
    SD = "Standard Deviation"
  ) %>%
  fmt_number(columns = c(Mean, Median, SD), decimals = 2) %>%
  tab_options(
    table.font.size = 12,
    heading.title.font.size = 16,
    heading.subtitle.font.size = 14
  )
```

---

# Inference

- If the **mean** and **median** differ substantially, this may suggest skewness in message volume.
- Compare between **Groups** to explore differences in texting behavior.
- An increase from **Baseline** to **Six_months** may indicate behavioral changes over time.
- Use standard deviation to understand variability within each subgroup.
